\documentclass[10pt]{article}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{amsopn}
\usepackage{latexsym}
\usepackage[utf8]{inputenc}
\usepackage{kpfonts}
\usepackage[T1]{fontenc}
\usepackage[polish,czech]{babel}
\usepackage[section]{algorithm}
\usepackage{latexsym}
\usepackage{color}
\usepackage{comment}
\usepackage{url}

\pagestyle{empty}

\begin{document}

\paragraph{Sebastian Siebertz.}
Post-doctoral research assistant at the Intitute of Informatics, Chair for Algorithm Engeneering, Humboldt-Universit\"at zu Berlin.

\paragraph{Address:}

 Institut f\"ur Informatik, Algorithm Engeneering\\
Humboldt-Universi\"at zu Berlin\\
Johann-von-Neumann-Haus\\
Rudower Chausse 25\\
D-12489 Berlin-Adlershof\\


\paragraph{Brief vita:}
Sebastian Siebertz received his PhD in 2015 from the Technical
University of Berlin. He was a postdoctoral researcher at
Technical University of Berlin from September 2015 until
September 2016. He was
a Marie Skłodowska-Curie (POLONEZ) fellow at the University
of Warsaw from October 2016 until September 2018. Since
October 2018 he is a postdoctoral researcher at Humboldt-Universit\"at Berlin.
His main research interests are in logical methods in computer
science and structural graph theory.

\paragraph{Selected publications:}
\begin{enumerate}
\item K.\ Eickmeyer, A.\ C.\ Giannopoulou, S.\ Kreutzer,
O.\ Kwon, M.\ Pilipczuk, R.\ Rabinovich, and S.\ Siebertz.
\emph{Neighborhood Complexity and Kernelization for Nowhere Dense Classes of Graphs.} 44th International Colloquium on Automata, Languages, and Programming, {ICALP} 2017.
\item J.~Gajarsk{\'{y}}, S.~Kreutzer, J.~Ne{\v s}et{\v r}il, P.~Ossona de Mendez, M.~Pilipczuk, S.~Siebertz and S.~Toru{\'{n}}czyk, \emph{First-order Interpretations of bounded expansion classes},
45th International Colloquium on Automata, Languages, and Programming, ICALP 2018.
\item  M.\ Grohe, S.\ Kreutzer, and S.\ Siebertz. \emph{Deciding first-order properties of nowhere dense graphs.} Journal of the ACM 64(3), pages 17:1--17:32, 2017.
\item S.\ Kreutzer, R.\ Rabinovich, and S.\ Siebertz. \emph{Polynomial Kernels and Wideness Properties of Nowhere Dense Graph Classes.} 28th Annual ACM-SIAM Symposium on Discrete Algorithms, SODA 2017.
\item M.~Pilipczuk, S.~Siebertz and S.~Toru\'nczyk, \emph{On the number of types in sparse graphs}.
33rd Annual ACM/IEEE Symposium on Logic in Computer Science, LICS 2018.
\end{enumerate}

\end{document}
